Breast Cancer Histopathological Images Segmentation and Classification using Vision Transformer

### 2021-04-25 paper/progress-1/01_rebuttal.tex

\documentclass[rebuttal]{cvpr}

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{0000} % Do not enter team number here
%\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\def\cvprPaperID{0000} % Do not enter team number here
\def\confYear{AI604 2021 spring}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{AI604 term project rebuttal / author response}  % **** Enter the paper title here

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
\section{Introduction}

After receiving paper reviews, authors should submit a rebuttal to address the reviewers' comments, which will be limited to a {\bf one page} PDF file.

Note that the author rebuttal is not visible to the reviewers. In other words, only TAs and the professor will read the rebuttal.

\subsection{Response length}

Author responses must be no longer than 1 page in length except for \textbf{References} and \textbf{Additional question}.

\subsection{Note}

Feel free to write the author response, and we recommend you not to spend your time so much on this. We aim to provide you a chance for improve your work by answering to the reviews.

%-------------------------------------------------------------------------
%\newpage

\section{Additional question}
Each team should answer the below questions for evaluating their reviewers.

\subsection{Question 1}
\begin{itemize}
    % 리뷰어 1 2 3 은 우리의 프로젝트 내용을 이해했다.
    \item The reviewer 1 understood our main idea and contributions. (Yes / Neutral / No)
    \item The reviewer 2 understood our main idea and contributions. (Yes / Neutral / No)
    \item ...
\end{itemize}

\subsection{Question 2}
\begin{itemize}
    % 리뷰어 1 2 3 은 우리에게 건설적인 피드백을 제공했다.
    \item The reviewer 1 provided constructive feedback or advice. (Yes / Neutral / No)
    \item The reviewer 2 provided constructive feedback or advice. (Yes / Neutral / No)
    \item ...
\end{itemize}

\subsection{Question 3}
\begin{itemize}
    % 기타 하고 싶은 말
    \item Other comments for TAs and the professor.
\end{itemize}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{reference}
}

\end{document}

### 2021-05-25 paper/progress-2/00_main.tex
Added dirty code for convolution free method.

Moreover, several studies such as \cite{Guofast2020} exclusively studies breast cancer WSI segmentation. In addition to WSI segmentation for cancer or disease detection, deep learning methods are being used for cancer classification from WSI\cite{XU2014591}, specifically for the  breast cancer. Most existing WSI segmentation methods are patch-based, therefore they require more additional post processing processing to combine the patches. Furthermore, some patches can be non-informative. Those methods commonly use deep CNN architectures which can slow the training process and difficult to train on computation resource limited devices. Transformer architectures can extract meaningful features from natural images. \\
\\
\textbf{Transformer.} Transformer architecture was first proposed by Vaswani et al.\cite{vaswani2017attention}. The essence of the architecture lies in the multi-headed self-attention (MHA) mechanism to learn the relationships among sequential tokens. MHA can be expressed as the following
\begin{equation}
\begin{split}
    \text{MHA}(Q, K,&\, V) =  \text{Concat}(A_1, \dots, A_h)W^O, \\[3mm]
    \text{where } A_i & := \text{Attention}(QW_i^Q, KW_i^K, VW_i^V) \\
        & = \text{softmax} \left( \frac{(QW_i^Q)(KW_i^K)^\top}{\sqrt{d_k}} \right) VW_i^V,
    \label{attention}
\end{split}
\end{equation}
where $Q, K, V$ are input query, key, value vectors, $W_i^Q, W_i^K \in \mathbb{R}^{d_{\text{model}}, d_k}, W_i^V \in \mathbb{R}^{d_{\text{model}}, d_v}$ are trainable query, key, value weights at the $i$-th attention head, $d_{\text{model}}$ the model dimension , $h$ number of heads, $d_k, d_v$ key and value dimension. By only using self-attention for processing sequential data, Transformer has shown state-of-the-art performance in numerous natural language processing tasks such as machine translation and sequence generation. Since its conception, Transformer architecture has been adapted to other domains.\\
\\
\textbf{Vision Transformer} Vision Transformer (ViT)\cite{dosovitskiy2020image} closely adapts Transformer Architecture for natural language processing for image classification. The concept consists of dividing the whole image into small chunks analogous to the sequence of tokens in machine translation. There are a few modifications from the original Transformer architecture for image domains:
\begin{enumerate}
    \item In order to circumvent the quadratic memory bottleneck, ViT takes in patches of images instead of pixels that are flattened then embedded via linear layer, essentially taking patches as tokens.
    \item Because of the sequential nature, ViT can take higher resolution inputs. However, position embeddings need to be interpolated for finetuning. This has been shown to be emprically effective in the subsequent downstream tasks. 
    \item As in the pretrained language model BERT, a classification token is appended before the patch tokens.
\end{enumerate}
Through extensive experiments, it has been proved that the ViT outperforms the existing state-of-the-art models in image classification tasks, thereby several variants of ViT have been proposed\cite{wu2021cvt,zhou2021deepvit}. Furthermore, ViT model can be a good feature extractor or encoder given that it has been trained on various type of datasets. Therefore, we apply this promising architecture to histological images segmentation and breast cancer classification.

% \begin{table}[ht]
% \centering
% \resizebox{\columnwidth}{!}{\begin{tabular}{c|c|c|c}
% \toprule
% Network & Salient Feature & Parameters & FLOP \\ \toprule
% AlexNet &  Deep & 61M & 1.5B \\ 
% VGG-16 & Fixed-size kernels & 138M & 19.6B \\
% Inception & Wide/parallel kernels & 43M & 2B \\
% ResNet-152 & Skip connection & 60M & 11B \\
% ViT & Transformer & 86M & \\
% \bottomrule
% \end{tabular}}
% \caption{Comparing number of trainable parameters and FLOPs according to model architectures. M stands for million and B billion.}
% \label{param}
% \end{table}

\section{Method}
\begin{figure}[ht]
\begin{center}
% \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{media/roi1.PNG}
\end{center}
   \caption{Block diagram of the model for WSIs segmentation and classification}
\label{fig:roi}
% \label{fig:onecol}
\end{figure}
Figure \ref{fig:roi} depicts the block diagram of the proposed method. Our approach can be decomposed into two main steps: first, we build a classification model based on ViT for breast cancer detection, then we build a model for WSI segmentation. We then combine those models to produce a multi tasks model. Another option we have is to apply domain adaptation to another organ segmentation dataset. 
\subsection{Classification}
\begin{figure}[ht]
\begin{center}
% \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{media/classify.png}
\end{center}
   \caption{Block diagram of the model for WSI classification}
\label{fig:cls}
% \label{fig:onecol}
\end{figure}
In this section we begin with the classification model. Our first goal is to use the ViT to accurately predict the class of breast cancer given a whole slide image as input. Previous works \cite{7780635} decompose the images into patches then fit those discriminative and non discriminative patches to a CNN one by one or resize to a lower resolution. We resize to an appropriate resolution (not necessarily lower for CNN) then use all patches of each image at once with ViT. Given an image $X^{W \times H \times C}$ with spatial resolution ${W \times H}$, ViT firstly divides it into $N=\frac{WH}{P^2}$ patches where $P$ is the patch size. This leads to each spatial dimension being a multiple of patch size. However, WSI do not always come in such manner. To deal with the design choice we propose to use rectangular patch size and also use padding strategy in case the sizes do not match. Although padding may have small effect on classification results, it is important for WSI segmentation. The new number of patches can be computed as in \ref{npatch}:
\begin{equation}
    N= \left\lceil\frac{H}{P_1}\right\rceil \cdot \left\lceil\frac{W}{P_2}\right\rceil,
    \label{npatch}
\end{equation}
where $P_1$ and $P_2$ are the spatial dimensions of a patch. Figure \ref{fig:cls} depicts the proposed architecture for cancer WSI classification.

\subsection{Segmentation}
\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.8\linewidth]{media/prev.png} 
\caption{Previous patch-based approach}
\label{fig:prev}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.8\linewidth]{media/prop.png}
\caption{Proposed segmentation architecture}
\label{fig:prop}
\end{subfigure}

\caption{WSI segmentation model diagram}
\label{fig:seg}
\end{figure}

If the image is too large we divide it into chunks of appropriate size then use those chunk as original image. Doing so we reduce the number on chunks to be treated sequentially compared to existing approaches which will produce large number of chunks then treats them sequentially \ref{fig:prev}. Most existing biomedical images segmentation approaches integrate UNet\cite{ronneberger2015unet} or a large CNN architecture. Our method also have the advantage to allow the insertion of part of an UNet module at the output of the transformer if needed.
Figure \ref{fig:prop} depicts the proposed segmentation approach. The model can produce a sequence of images at once if the ground true is a sequence of images like in machine translation. Also, we can bypass the post processing part and produce accurate segmentation mask. The reconstruction head in the Figure \ref{fig:prop} can be a set of convolution layers to produce the appropriate final shape for the output model.

\subsection{Evaluation metric}
To evaluate the proposed method, we use two different metrics for prediction and segmentation. The prediction level metric consists of: \textit{F1}, \textit{precision}, and the \textit{sensitivity} as defined in \ref{metric1}

\begin{equation}
\begin{split}
   \text{F1} & = \frac{2TP}{2TP+FP+ FN}, \\
   \text{Sensitivity} & = \frac{TP}{TP+FN}, \\
   \text{Precision} & = \frac{TP}{TP+FP} ,
   \label{metric1}
\end{split}
\end{equation}
where $TP$, $FP$ and $FN$ are true positive, false positive and false negative, respectively.
For segmentation level metric, the intersection of Union metric: Jacard index or DICE's coefficient $D$ defined in equation \ref{metric2},
\begin{equation}
    D(\hat{Y},Y)=2\frac{|\hat{Y} \cap Y|}{|\hat{Y} \cup Y|},
    \label{metric2}
\end{equation}
where $\hat{Y}$ and $Y$ are the predicted region from the model and the ground truth region of input image, respectively.

\section{Preliminary experiments}
At this stage of our work we evaluate the proposed model on breast cancer data compared to some recent works. The proposed classifier was trained using an RTX2080ti and 4 RTX TitanX GPUs  with TensorFlow and PyTorch library. Most of the training was done on RTX2080ti desktop.
\subsection{Data}
Breast Cancer Histopathological Image Classification (BreakHis) dataset \cite{BENHAMMOU20209} was used to assess the performance. This dataset is composed of 9,109 microscopic images of breast tumor tissue collected from 82 patients using different magnifying factors (40X, 100X, 200X, and 400X). It contains 2,480  benign and 5,429 malignant samples ($700\times 450$ pixels, 3-channel RGB, 8-bit depth in each channel, PNG format). Benign and malignant subtypes \ref{TAB1} are also provided which makes this dataset an good choice for comparing models performance. According to the dataset, there are four histological distinct types of benign breast tumors (B): adenosis (A), fibroadenoma (F), phyllodes tumor (PT), and tubular adenona (TA) and four malignant breast tumors (M): carcinoma (DC), lobular carcinoma (LC), mucinous carcinoma (MC) and papillary carcinoma (PC).

@@

% \begin{table}[ht]
% \centering
% \begin{tabular}{c|c|cccc|c|c} %{llllllll}
% \hline
% \multirow{2}{*}{Class}   & \multirow{2}{*}{Type } & \multicolumn{4}{l|}{Magnification} & \multirow{2}{*}{Total} & \multirow{2}{*}{Sum}    \\
%                           &   & 40X  & 100X & 200X & 400X & & \\
%                           \hline
% \multirow{4}{*}{  B } & A & 114  & 113  & 111  & 106 & 444 & \multirow{4}{*}{2,480}  \\
%                           & F  & 253  & 260  & 264  & 237 & 1014 & \\
%                           & TA & 109 & 121  & 108  & 115 & 453 & \\
%                           & PT & 149  & 150  & 140  & 130 & 569 & \\
%                           \hline
% \multirow{4}{*}{M} & DC & 864  & 903  & 896  & 788 & 3451  & \multirow{4}{*}{5,429}  \\
%                           & LC & 156  & 170  & 163  & 137  & 626 & \\
%                           & MC & 205  & 222  & 196  & 169 & 792 & \\
%                           & PC & 145  & 142  & 135  & 138 & 560 &  \\
%                           \hline
% \multicolumn{2}{|l|}{Total}& 1995 & 2081 & 2013 & 1820 & 7909 & 7,909\\
% \hline
% \end{tabularx}
% \caption{BreaKHis dataset description}
% \label{TAB1}
% \end{table}

\subsection{Breast cancer classification using BreakHis dataset}
We consider two approaches to evaluate our model which are magnification specific binary(MSB) classification  and  multi classes classification. For the binary classification, we only consider benign and malignant to be the classes. For multi classes classification, we only consider the subclasses. The results of the experiment for binary classification are showed in Table \ref{result1}. In this experiment, for fair comparison we consider two state-of-the-art methods\cite{Xie2019,8353225} With similar setting on the same dataset.  Data augmentation techniques  consisting to increase the number of samples where not used. Also, we do not use any pretrained techniques.

\begin{table}[ht]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{cc|cccc|c}
\toprule
\multirow{2}{*}{Works} & \multirow{2}{*}{Year} & \multicolumn{4}{c|}{Magnification} & \multirow{2}{*}{Mean} \\
 & & 40X & 100X & 200X & 400X & \\ \midrule
\cite{Xie2019} & '19 & 97.90 & 96.88 & 96.88 & 96.88 & 97.13 \\
Ours & '21 & \textbf{98.36} & 96.19 & 95.86 & \textbf{97.98} & 97.10 \\
VGG\cite{8353225} & '18 & 96.82 & 96.96 & 96.36 & 95.97 & 96.52 \\
Ens\cite{8353225} & '18 & 98.33 & \textbf{97.12} & \textbf{97.85} & 96.15 & \textbf{97.36} \\
\bottomrule
\end{tabular}}
\caption{Comparison of MSB classification results. VGG stands for VGG-16 architecture and Ens stands for Ensemble method proposed in \cite{8353225}. Boldface means the maximum within the respective column.}
\label{result1}
\end{table}

In this preliminary experiment we trained the proposed model from scratch. For comparison, the BreakHis dataset is randomly split into 70\% for training and 30\% for testing. We only applied horizontal flip and normalization. All methods are within in the same range although those method used complex convolution architectures. Based on the results presented in \cite{BENHAMMOU20209}, our approach achieves competitive compared to existing breast cancer classification method tested on BreakHis. However, further improvement are still needed. Therefore, we will perform further experiments, parameter tuning to outperform the state-of-the-art on BreakHis and also move to tissues segmentation. The results presented in this paper are preliminary results and the final results may different significantly. Therefore we have not yet include the evaluation metrics results.

\subsection{Additional experiments with pretrained ViT}
Some additional experiments were conducted with publicly available pretrained ViT model. This was to investigate the effect of pretraining on a drastically different image domains to lay basis for further investigation. For this model, the image was resized to 224-by-224 for preliminary experiment. However, this part will be modified in the coming end-to-end implementation. For experiment, a base model with patch size 16 and input size 224 was used. The experiments can be viewed \href{https://wandb.ai/jkgrad/vit-domain-adapt?workspace=user-jkgrad}{here}.

The results are provided Table \ref{result2}. As expected, it was found that pretrained models can absolutely leverage information obtained during pretraining. Even without data augmentation, the model already achieves the state-of-the-art result. When coupled with data augmentation, the model performance increases at low magnification, leading to the mean accuracy across magnifications levels of 98.52, which is remarkable. The effect of higher input resolution for downstream task, however, still needs to be investigated.

\begin{table}[ht]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{c|cccc|c}
\toprule
\multirow{2}{*}{Works} & \multicolumn{4}{c|}{Magnification} & \multirow{2}{*}{Mean} \\
 & 40X & 100X & 200X & 400X & \\ \midrule
PT    & 97.66 & \textbf{98.72} & \textbf{98.01} & \textbf{98.35} & 98.19 \\
PT+AG & \textbf{99.00} & \textbf{98.72} & \textbf{98.01} & \textbf{98.35} & \textbf{98.52} \\
\bottomrule
\end{tabular}}
\caption{Experiments on pretraining. All models were trained under the same hyperparameters. PT stands for pretraining and AG stands for augmentation.}
\label{result2}
\end{table}

\section{Conclusion}
In this stage of work  we proposed a classification model for histological images using ViT architecture. The proposed method is fast to train and achieved good performance compared to previous works. In the coming stage, we will evaluate the model performance using the proposed metrics. Also, we will conduct more classification experiments to include the multi classes classification results. After that we will apply the proposed architecture for WSI segmentation. Finally we will investigate transfer learning and domain adaptation.

{\small
\bibliographystyle{ieee_fullname}
\bibliography{reference}
}

>>>>

Moreover, several studies such as \cite{Guofast2020} exclusively studies breast cancer WSI segmentation. In addition to WSI segmentation for cancer or disease detection, deep learning methods are being used for cancer classification from WSI\cite{XU2014591}, specifically for the  breast cancer. Most existing WSI segmentation methods are patch-based, therefore they require more additional post processing processing to combine the patches. Furthermore, some patches can be non-informative. Those methods commonly use deep CNN architectures which can slow the training process and difficult to train on computation resource limited devices. Transformer architectures can extract meaningful features from natural images. \\
\\
\textbf{Transformer.} Transformer architecture was first proposed by Vaswani et al.\cite{vaswani2017attention}. The essence of the architecture lies in the multi-headed self-attention (MHA) mechanism to learn the relationships among sequential tokens. MHA can be expressed as the following
\begin{equation}
\begin{split}
    \text{MHA}(Q, K,&\, V) =  \text{Concat}(A_1, \dots, A_h)W^O, \\[3mm]
    \text{where } A_i & := \text{Attention}(QW_i^Q, KW_i^K, VW_i^V) \\
        & = \text{softmax} \left( \frac{(QW_i^Q)(KW_i^K)^\top}{\sqrt{d_k}} \right) VW_i^V,
    \label{attention}
\end{split}
\end{equation}
where $Q, K, V$ are input query, key, value vectors, $W_i^Q, W_i^K \in \mathbb{R}^{d_{\text{model}}, d_k}, W_i^V \in \mathbb{R}^{d_{\text{model}}, d_v}$ are trainable query, key, value weights at the $i$-th attention head, $d_{\text{model}}$ the model dimension , $h$ number of heads, $d_k, d_v$ key and value dimension. By only using self-attention for processing sequential data, Transformer has shown state-of-the-art performance in numerous natural language processing tasks such as machine translation and sequence generation. Since its conception, Transformer architecture has been adapted to other domains.\\
\\
\textbf{Vision Transformer} Vision Transformer (ViT)\cite{dosovitskiy2020image} closely adapts Transformer Architecture from natural language processing for image classification. The concept consists of dividing the whole image into small chunks analogous to the sequence of language tokens. There are a few modifications from the original Transformer architecture for image domains:
\begin{enumerate}
    \item In order to circumvent the quadratic memory bottleneck, ViT takes in patches of images instead of pixels that are flattened then embedded via linear layer, essentially taking patches as tokens.
    \item Because of the sequential nature, ViT can take higher resolution inputs. However, position embeddings need to be interpolated for finetuning. This has been shown to be emprically effective in the subsequent downstream tasks. 
    \item As in the pretrained language model BERT, a classification token is appended before the patch tokens.
\end{enumerate}
Through extensive experiments, it has been proved that the ViT outperforms the existing state-of-the-art models in image classification tasks, thereby several variants of ViT have been proposed\cite{wu2021cvt,zhou2021deepvit}. Furthermore, ViT model can be a good feature extractor or encoder given that it has been trained on various type of datasets. Therefore, we apply this promising architecture to histological images segmentation and breast cancer classification.

% \begin{table}[ht]
%     \centering
%     \begin{tabular}{cccc}
%         \toprule
%         Models & Image size & \# Params & FLOPs  \\ \midrule
%         \multicolumn{4}{c}{Convolution} \\ \midrule
%         Resnet-101 & $224^2$ & 45M & \\
%         EffNet-B3 & $300^2$ & 12M & 1.8G \\
%         EffNet-B4 & $380^2$ & 19M & 4.2G \\ \midrule
%         \multicolumn{4}{c}{Transformers} \\ \midrule
%         ViT-B/16 & $384^2$ & 86M & 55.4G \\
%         ViT-L/16 & $384^2$ & 307M & 190.7G \\
%         DeiT-B   & $224^2$ & 22M & 4.6G \\
%         DeiT-B   & $384^2$ & 86M & 17.5G \\ \bottomrule
%     \end{tabular}
%     \caption{Computation and parameter comparison of different image model backbones on ImageNet. The table was adapted from \cite{touvron2021training}.}
%     \label{computation_comparison}
% \end{table}

\section{Method}
\begin{figure}[ht]
\begin{center}
% \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{media/roi1-new.png}
\end{center}
   \caption{Block diagram of the model for WSIs segmentation and classification}
\label{fig:roi}
% \label{fig:onecol}
\end{figure}
Figure \ref{fig:roi} depicts the block diagram of the proposed method. Our approach can be decomposed into two main steps: first, we build a classification model based on ViT for breast cancer detection, then we build a model for WSI segmentation. We then combine those models to produce a multi tasks model. Another option we have is to apply domain adaptation to another organ segmentation dataset. 
\subsection{Classification}
\begin{figure}[ht]
\begin{center}
% \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{media/classify-new.png}
\end{center}
   \caption{Block diagram of the model for WSI classification}
\label{fig:cls}
% \label{fig:onecol}
\end{figure}
In this section we begin with the classification model. Our first goal is to use the ViT to accurately predict the class of breast cancer given a whole slide image as input. Previous works \cite{7780635} decompose the images into patches then fit those discriminative and non discriminative patches to a CNN one by one or resize to a lower resolution. We resize to an appropriate resolution (not necessarily lower for CNN) then use all patches of each image at once with ViT. Given an image $X^{H \times W \times C}$ with spatial resolution ${H \times W}$, ViT firstly divides it into $N = \frac{HW}{P^2}$ patches where $P$ is the patch size. This leads to each spatial dimension being a multiple of patch size. However, WSI do not always come in such manner. To deal with the design choice we propose to use rectangular patch size and also use padding strategy in case the sizes do not match. Although padding may have small effect on classification results, it is important for WSI segmentation. The new number of patches can be computed as in \ref{npatch}:
\begin{equation}
    N= \left\lceil\frac{H}{P_1}\right\rceil \cdot \left\lceil\frac{W}{P_2}\right\rceil,
    \label{npatch}
\end{equation}
where $P_1$ and $P_2$ are the spatial dimensions of a patch. Figure \ref{fig:cls} depicts the proposed architecture for cancer WSI classification.

\subsection{Segmentation}
\begin{figure}[h]
    \centering
    \begin{subfigure}{0.5\textwidth}
        \centering
        \resizebox{0.8\columnwidth}{!}{\includegraphics[width=0.8\linewidth]{media/prev-new.png}}
        \caption{Previous patch-based approach}
        \label{fig:prev}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \centering
        \resizebox{0.8\columnwidth}{!}{\includegraphics[width=0.8\linewidth]{media/prop-new.png}}
        \caption{Proposed segmentation architecture}
        \label{fig:prop}
    \end{subfigure}
\caption{WSI segmentation model diagram}
\label{fig:seg}
\end{figure}

If the image is too large we divide it into chunks of appropriate size then use those chunk as original image. Doing so we reduce the number on chunks to be treated sequentially compared to existing approaches which will produce large number of chunks then treats them sequentially \ref{fig:prev}. Most existing biomedical images segmentation approaches integrate UNet\cite{ronneberger2015unet} or a large CNN backbone that can only take fixed input sizes. Our method also have the advantage to allow the insertion of part of an UNet module at the output of the transformer if needed.

Figure \ref{fig:prop} depicts the proposed segmentation approach. The model can produce a sequence of images at once if the ground true is a sequence of images like in machine translation. Also, we can bypass the post processing part and produce accurate segmentation mask. The reconstruction head in the Figure \ref{fig:prop} can be a set of convolution layers to produce the appropriate final shape for the output model. \\
The input image to our segmentation network is a 2D image of size $(H \times W \times C)$ where $H=W$ are the image height and width, and $C$ the number of channels (3 for RGB images).

The input image is then partitioned to a sequence of patches of size $(P\times P\times C)$. These patches are passed through the transformer which outputs the embedding matrix of shape $(B, N, d)$,  where $B$ is the batch size, $N$ the number of patches and $d$ the embedding dimension. For reconstructing the prediction mask, we can add a multi-layer perceptron to transform that embedding into a matrix of shape $(B, cHW)$ and reshape to $(B, c, H, W)$ where $c$ is the number of classes. Such an approach requires a lot of memory resources for large images. Another way to reconstruct the segmentation mask is to directly reshape the embedding matrix into $(B, N, d_1, d_2)$ where $d = d_1\times d_2$, then apply transpose convolution as in UNet architecture. This approaches reduces largely the numbers of convolution layers while playing the role of the encoder in UNet-like architecture.

\subsection{Evaluation metric}
To evaluate the proposed method, we use the following metrics for each tasks.

\subsubsection{Classification}
\begin{enumerate}
    \item[\textbullet] Accuracy: a naive multi-class classification metric. 
    \item[\textbullet] F-1 score: given TP, FP, and FN are true positive, false positive, and false negative, F-1 score can be computed as 
    \begin{equation}
         \text{F-1} = \frac{2\text{TP}}{2\text{TP}+\text{FP}+ \text{FN}}
         \label{f1-metric}
    \end{equation}
\end{enumerate}

\subsubsection{Segmentation}
\begin{enumerate}
    \item[\textbullet] Pixel accuracy: a naive pixel-wise accuracy metric. 
    \item[\textbullet] Intersection over Union (IoU/Jaccard): IoU, also known as Jaccard index, can be computed in a similar manner to \ref{f1-metric}, but without the constant:
    \begin{equation}
        \text{Dice} = \frac{\text{TP}}{\text{TP}+\text{FP}+ \text{FN}}
        \label{dice-score-metric}
    \end{equation}
    \item[\textbullet] Dice coefficient (Dice/F-1): Dice coefficient is defined to be the following where $\hat{Y}$ and $Y$ are the predicted region from the model and the ground truth region of input image.
    \begin{equation}
        \text{Dice} (\hat{Y},Y)=2\frac{|\hat{Y} \cap Y|}{|\hat{Y} \cup Y|},
        \label{dice-score}
    \end{equation}
    This is equivalent to F-1 score in the context of segmentation, which can be computed by \ref{f1-metric}.
\end{enumerate}

% \begin{equation}
% \begin{split}
%   \text{F1} & = \frac{2\text{TP}}{2\text{TP}+\text{FP}+ \text{FN}}, \\
%   \text{Sensitivity} & = \frac{TP}{TP+FN}, \\
%   \text{Precision} & = \frac{TP}{TP+FP} ,
%   \label{metric1}
% \end{split}
% \end{equation}

\section{Experiments}
\subsection{Classification}
\subsubsection{Dataset description}
Breast Cancer Histopathological Image Classification (BreakHis) dataset \cite{BENHAMMOU20209} was used to assess the performance. This dataset is composed of 9,109 microscopic images of breast tumor tissue collected from 82 patients using different magnifying factors (40X, 100X, 200X, and 400X). It contains 2,480  benign and 5,429 malignant samples ($700\times 450$ pixels, 3-channel RGB, 8-bit depth in each channel, PNG format). Benign and malignant subtypes \ref{TAB1} are also provided which makes this dataset an good choice for comparing models performance. According to the dataset, there are four histological distinct types of benign breast tumors (B): adenosis (A), fibroadenoma (F), phyllodes tumor (PT), and tubular adenona (TA) and four malignant breast tumors (M): carcinoma (DC), lobular carcinoma (LC), mucinous carcinoma (MC) and papillary carcinoma (PC).

@@

\subsubsection{Experiment setup}
We consider two approaches to evaluate our model which are magnification specific binary(MSB) classification  and  magnification specific multi classes classification (MSM). For the binary classification, we only consider benign and malignant to be the classes. For multi classes classification, we only consider the subclasses. The results of the experiment for binary classification are showed in Table \ref{result1}. In this experiment, we consider two state-of-the-art methods \cite{Xie2019,8353225} on the same dataset.  Data augmentation techniques were not used. Also, we do not use any pretrained models.

We  train the model for both binary and multi-class classification in a multi-label classification manner. In this experiment we train the proposed model from scratch with patch size $(20 \times 50)$ and input size $(460\times 700 \times 3)$. For comparison, the BreakHis dataset is randomly split into 70\% for training and 30\% for testing for each magnification factor. We only consider the methods that preprocess the dataset in the same manner as we do. Table \ref{result0} presents the experimentation results.

\subsubsection{Result}
The model achieves good performance on the disease classes compared to the subclasses. However, the performance of the subclasses can be improved if they are trained separately. The performance on the subclasses is comparable to existing methods \cite{BENHAMMOU20209,Xie2019}. Based on the results presented in \cite{BENHAMMOU20209}, our approach achieves competitive results \ref{result1} compared to existing breast cancer classification method tested on BreakHis. However, further improvement are still needed for the multilabels classification. Therefore, we will perform further experiments with non combined training, parameter tuning to outperform the state-of-the-art on BreakHis and move to tissues segmentation. The results presented in this paper are preliminary results and the final results may different significantly.

\begin{table}[ht]
\centering
\resizebox{\columnwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|} 
\hline
\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Classification\end{tabular}} & \multirow{2}{*}{criteria} & \multicolumn{4}{l|}{Magnification}  \\ 
\cline{3-6} &  & 40X   & 100X  & 200X   & 400 \\ 
\hline
\multirow{3}{*}{MSB  } & F-1 micro & 99.00 & 98.00 & 100.00 & 99.00  \\ 
\cline{2-6}& F-1 macro & 98.88 & 97.68 & 99.45  & 99.37\\ 
\cline{2-6}& Acc. & 99.00 & 98.08 & 99.50  & 99.45 \\ 
\hline
\multirow{3}{*}{MSM} & F-1 micro & 89.00    & 90.00   & 90.00     &90.00 \\ 
\cline{2-6} & F-1 macro & 87.10 & 86.99 &86.33    & 88.13         \\ 
\cline{2-6} & Acc.  & 89.48  & 89.76    & 90.23    & 89.74 \\
\hline
\end{tabular}}
\caption{Evaluation results on Breakhis dataset.}
\label{result0}
\end{table}

\begin{table}[ht]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{cc|cccc|c}
\toprule
\multirow{2}{*}{Works} & \multirow{2}{*}{Year} & \multicolumn{4}{c|}{Magnification} & \multirow{2}{*}{Mean} \\
 & & 40X & 100X & 200X & 400X & \\ \midrule
\cite{Xie2019} & '19 & 97.90 & 96.88 & 96.88 & 96.88 & 97.13 \\
Ours & '21 & \textbf{99.00} & \textbf{98.08} & \textbf{99.50} & \textbf{99.45} & \textbf{99.01} \\
VGG\cite{8353225} & '18 & 96.82 & 96.96 & 96.36 & 95.97 & 96.52 \\
Ens\cite{8353225} & '18 & 98.33 & 97.12 & 97.85 & 96.15 & 97.36 \\
\bottomrule
\end{tabular}}
\caption{Comparison of MSB classification results. VGG stands for VGG-16 architecture and Ens stands for Ensemble method proposed in \cite{8353225}. Boldface means the maximum within the respective column.}
\label{result1}
\end{table}

% \subsection{Additional experiments with pretrained ViT}
% Some additional experiments were conducted with publicly available pretrained ViT model. This was to investigate the effect of pretraining on a drastically different image domains to lay basis for further investigation. For this model, the image was resized to 224-by-224 for preliminary experiment. However, this part will be modified in the coming end-to-end implementation. For experiment, a base model with patch size 16 and input size 224 was used. The experiments can be viewed \href{https://wandb.ai/jkgrad/vit-domain-adapt?workspace=user-jkgrad}{here}.

% The results are provided Table \ref{result2}. As expected, it was found that pretrained models can absolutely leverage information obtained during pretraining. Even without data augmentation, the model already achieves the state-of-the-art result. When coupled with data augmentation, the model performance increases at low magnification, leading to the mean accuracy across magnifications levels of 98.52, which is remarkable. The effect of higher input resolution for downstream task, however, still needs to be investigated.

% \begin{table}[ht]
% \centering
% \resizebox{\columnwidth}{!}{\begin{tabular}{c|cccc|c}
% \toprule
% \multirow{2}{*}{Works} & \multicolumn{4}{c|}{Magnification} & \multirow{2}{*}{Mean} \\
%  & 40X & 100X & 200X & 400X & \\ \midrule
% PT    & 97.66 & \textbf{98.72} & \textbf{98.01} & \textbf{98.35} & 98.19 \\
% PT+AG & \textbf{99.00} & \textbf{98.72} & \textbf{98.01} & \textbf{98.35} & \textbf{98.52} \\
% \bottomrule
% \end{tabular}}
% \caption{Experiments on pretraining. All models were trained under the same hyperparameters. PT stands for pretraining and AG stands for augmentation.}
% \label{result2}
% \end{table}

\subsubsection{Dataset description}
We evaluate the model performance on WSIs segmentation using crowdsourced dataset presented in \cite{10.1093/bioinformatics/btz083}. This dataset consists of 151 annotated WSIs and corresponding masks with each image's dimension larger than 1024 and each image with different dimensions. For more conveninece, we partition all the images into patches of size $1024 \times 1024$. After patching, there are a total number of 4235 pairs of RGB image and mask which were split into train-test set using the same ratio as the classification task.

The networks were trained to map pixels into 22 different classes including the outside region of interest (``don't care" class). The RGB images and their corresponding masks were 0.23 microns per pixel where each mask pixel values encode region class membership. We first implement some existing baseline models including U-Net. However, training such a model with input size $(1024 \times 1024)$ requires a lots of computation resources. Therefore, we repartition the images into patches of size $(512\times 512)$. We also decide to train the proposed method with same input size and patch size of $(128\times 128)$ using transpose convolution reconstruction head. In this setting the model performs 1 epoch in every 3 minutes during training.

\subsection{Comparison with state-of-the-arts}
We will compare our model with 3 state-of-the-art models: (1) UNet \cite{ronneberger2015unet}, (2) UNet with pretrained MobileNetV2 as encoder (UNet-MV2) and (3) TransUNet \cite{chen2021transunet}. For all experiments, we applied simple augmentations such as horizontol flip and normalization. It took around 3 hours each to train UNet and UNet-MV2 models on TeslaV100 GPU for 15 epochs. The segmentation results and model accuracies are shown in Figure \ref{fig:seg} and Table. We have yet to include the results from TransUNet due to its long training time (1 hour per epoch) for better convergence.
